\documentclass[aspectratio=169,11pt]{beamer}

\usetheme{Madrid}
\usecolortheme{default}
\usefonttheme{professionalfonts}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}[frame number]

\usepackage{amsmath, amssymb, booktabs}

\newcommand{\E}{\mathbb{E}}
\newcommand{\1}{\mathbf{1}}
\newcommand{\plim}{\operatorname{plim}}

\title{Extraction, Classification, Embeddings, And Validation}
\subtitle{Empirical Methods --- Week 13}
\author{Teaching deck draft}
\date{}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Week 13 question}
\begin{itemize}
    \item Once unstructured data become variables, how do we know what they measure?
    \item Validation is part of identification when the constructed variable enters the economics.
    \item The goal is not another extraction workflow.
    \item The goal is to defend benchmarks, diagnostics, robustness, transportability, and downstream interpretation.
\end{itemize}

\medskip
Text, images, audio, video, and multimodal records all need the same discipline.
\end{frame}

\begin{frame}{Validation as identification}
\small
\[
\widehat M_i = g(U_i;\widehat\theta)
\]

\begin{tabular}{p{0.24\linewidth} p{0.66\linewidth}}
\toprule
Layer & Question \\
\midrule
Construct & Does the measure match the economic object? \\
Benchmark & Do labels or ground truth represent that object? \\
Predictive & Does the model recover labels out of sample? \\
Subgroup & Does performance vary by group, place, time, or modality? \\
Downstream & Do remaining errors change descriptive or causal claims? \\
\bottomrule
\end{tabular}

\medskip
A strong downstream design cannot rescue a poorly defined measure.
\end{frame}

\begin{frame}{Annotation and benchmark design}
\small
\begin{tabular}{p{0.27\linewidth} p{0.31\linewidth} p{0.31\linewidth}}
\toprule
Choice & What to specify & Why it matters \\
\midrule
Unit & document, image, clip, response & defines the estimand and error structure \\
Label & construct vs proxy & determines what the model learns \\
Annotators & experts, RAs, crowd, subjects & changes interpretation and quality \\
Rubric & rules, examples, edge cases & makes disagreement diagnosable \\
Adjudication & vote, expert review, tie-break & shapes benchmark stability \\
Holdout & untouched labeled sample & prevents tuning to the benchmark \\
\bottomrule
\end{tabular}

\medskip
\[
\kappa=\frac{p_o-p_e}{1-p_e}
\]

Agreement is useful, but agreement on the wrong construct is still wrong.
\end{frame}

\begin{frame}{Precision, recall, and F1}
\small
\[
\text{Precision}=\frac{TP}{TP+FP},
\qquad
\text{Recall}=\frac{TP}{TP+FN}
\]

\[
F_1
=
2\cdot
\frac{\text{Precision}\cdot\text{Recall}}
{\text{Precision}+\text{Recall}}
\]

\begin{itemize}
    \item Precision matters when false positives are costly.
    \item Recall matters when missed positives are costly.
    \item F1 balances the two, but it is not a policy loss function.
    \item The right threshold depends on the research use.
\end{itemize}
\end{frame}

\begin{frame}{Calibration}
\small
\[
\E[Y\mid \widehat p=p]=p
\]

\begin{itemize}
    \item A model can rank cases well and still give bad probabilities.
    \item Calibration matters for prevalence, targeting, thresholding, welfare calculations, and cost-benefit analysis.
    \item Check calibration overall and by subgroup.
    \item Recalibration may fix probabilities without fixing construct validity.
\end{itemize}

\medskip
In applied economics, a score is often interpreted. Calibration tells us whether that interpretation is defensible.
\end{frame}

\begin{frame}{Subgroup performance and fairness}
\small
\[
\Pr(\widehat Y_i\neq Y_i\mid G_i=g)
\]

\begin{tabular}{p{0.23\linewidth} p{0.33\linewidth} p{0.33\linewidth}}
\toprule
Setting & Failure mode & Economic consequence \\
\midrule
Job postings & lower recall for service work & understated remote-work access \\
Satellite images & weak transfer to dense informal areas & distorted spatial inequality \\
Audio & device or identity shortcuts & tone measure confounds speaker traits \\
Video & high labeling burden and context loss & interaction score lacks construct validity \\
\bottomrule
\end{tabular}

\medskip
Fairness diagnostics are also identification diagnostics when group differences enter the claim.
\end{frame}

\begin{frame}{Drift and transportability}
\small
\begin{itemize}
    \item Vocabulary changes across firms, platforms, occupations, and time.
    \item Images change with sensors, seasons, resolution, and geography.
    \item Audio changes with microphones, rooms, speakers, and recording norms.
    \item Video changes with platform, consent, camera angle, and interaction setting.
\end{itemize}

\medskip
Transport checks:
\begin{itemize}
    \item hold out time, place, firm, occupation, annotator, or sensor;
    \item relabel in the target domain;
    \item report what breaks before using the measure downstream.
\end{itemize}
\end{frame}

\begin{frame}{Embedding similarity and external validation}
\small
\[
S_i(c)=\frac{e_i'c}{\|e_i\|\|c\|}
\]

\begin{itemize}
    \item Embeddings generalize beyond exact words or visual patterns.
    \item Similarity is not automatically exposure, feasibility, poverty, tone, or quality.
    \item Show nearest-neighbor audits, false positives, and false negatives.
    \item Compare to independent benchmarks not used in training.
    \item Test whether the same interpretation survives across groups and domains.
\end{itemize}
\end{frame}

\begin{frame}{Measurement error in downstream economics}
\small
\[
\widetilde X_i=X_i+\nu_i
\]

\[
\plim \widehat\beta_{\widetilde X}
=
\beta
+
\frac{\operatorname{Cov}(\widetilde X_i,\varepsilon_i-\beta\nu_i)}
{\operatorname{Var}(\widetilde X_i)}
\]

\begin{itemize}
    \item Classical error can attenuate simple slopes.
    \item Differential error can change sign, subgroup gaps, and event-study dynamics.
    \item If measurement error is correlated with treatment, time, group, or outcomes, the economics changes.
    \item Robustness should include downstream estimates under alternative validated measures.
\end{itemize}
\end{frame}

\begin{frame}{Validation examples from papers}
\small
\begin{tabular}{p{0.26\linewidth} p{0.27\linewidth} p{0.36\linewidth}}
\toprule
Paper setting & Modality & Validation lesson \\
\midrule
Hansen et al. & job-posting text & labels and model checks support remote-work comparisons \\
Haaland et al. & open-ended responses & coding, prompts, and human review define the object \\
Jean et al. & satellite imagery & image proxies need survey ground truth \\
Gorodnichenko et al. & audio & voice features need economic interpretation beyond acoustics \\
Blattman et al. & survey validation & qualitative evidence can reveal benchmark error \\
\bottomrule
\end{tabular}

\medskip
The shared standard: show what the measure captures and where it fails.
\end{frame}

\begin{frame}{Research Lab design}
\begin{columns}[T,onlytextwidth]
\begin{column}{0.32\linewidth}
\textbf{Reproduce}
\begin{itemize}
    \item labeled benchmark
    \item rater agreement
    \item dictionary baseline
    \item embedding similarity
    \item supervised score
\end{itemize}
\end{column}
\begin{column}{0.32\linewidth}
\textbf{Diagnose}
\begin{itemize}
    \item confusion matrix
    \item calibration bins
    \item subgroup metrics
    \item error audits
    \item downstream bias
\end{itemize}
\end{column}
\begin{column}{0.32\linewidth}
\textbf{Transfer}
\begin{itemize}
    \item new domain text
    \item vocabulary shift
    \item image proxy
    \item transport metrics
    \item redesign memo
\end{itemize}
\end{column}
\end{columns}

\medskip
The lab validates a constructed variable before interpreting it economically.
\end{frame}

\begin{frame}{Takeaways}
\begin{itemize}
    \item Define the construct and benchmark before choosing the model.
    \item Report metrics that match the research use, not only headline accuracy.
    \item Audit subgroup performance, calibration, drift, and transfer.
    \item Treat human review and external validation as core evidence.
    \item Explain how residual measurement error affects downstream economics.
\end{itemize}
\end{frame}

\end{document}
